\chapter{Technical Background}

In this chapter, I provide some technical introduction to relevant topics. Further information is available in the linked resources.

\section{Definitions}

\subsection{Filesystem}
\label{sec:fs}

\Glspl{storagemedia} takes many forms. There are the traditional \gls{hdd} and newer \gls{ssd}. Both are built based on different assumptions and using different technologies. There exist more \glspl{storagemedia} like \gls{cd}, \gls{dvd}, \gls{usb} flash drives and more. \Glspl{storagemedia} operates on blocks of data, which are typicaly 512 bytes or 4 \gls{kb} in size. \cite{bruce:imaging}

This creates a problem when an \gls{os} want to use a \gls{storagemedia} to store data. The \gls{os} generally wants to store files which can be any size from some bytes up to several \gls{gb}. \Glspl{fs} were created that the \gls{os} does not need to be concerned with data blocks. They create a layer of abstraction for the \gls{storagemedia} and offer the \gls{os} a simple and standard way to access files. Another benefit of \glspl{fs} is that they store \gls{metadata} such as timestamps, permissions, and attributes for each file. This data helps the \gls{os} to identify relevant data.

The \gls{fs} has it's own datastructure. It uses so called inodes to store the \gls{metadata} and uses the blocks of the \gls{storagemedia} to store the files themselves. Which \gls{metadata} is available heavily depends on the \gls{fs} used. There are some standard \gls{metadata}, that almost all \gls{fs} support, but there are also more \gls{metadata} that only certain \gls{fs} offer \cite{bruce:imaging}.

\subsection{Cryptographic Hashing Function}
\label{sec:hashing}

A hash function is, in general, a function which takes inputs of unspecified size and generates an output of a fixed size. Hash functions are used widely in computer science. One examples are programming and databases where they are used to access certain data within a data structure easily. By design, hash functions have \glspl{collision}, meaning that multiple inputs generate the same output. This statement must be true if you consider the unlimited input size and limited output size. If there are more inputs than outputs, there must be at least one output value that is assigned to multiple inputs. For data storage and other use cases this is not a huge problem because collisions can be handled, and if the hash function has good distribution, collisions are unlikely. Additionally, in many systems, a hashing function with weak collision resistance is deliberately chosen because it is usually faster to execute. \cite{hash:noncrypto, hash:slow}

In a cryptographic context, this tradeoff cannot be taken. Two big factors play into why not. Firstly, in cryptography, hashes are often used as an assurance that the content of some data has not changed. If collisions are easy to find, the data can be altered in ways that result in the same hash, meaning that the hash no longer fulfills the use case. Additionally, monetary incentive can be big. If a collision can be provoked, even if challenging, data can be changed. This data could be a legal document or a bank transaction, neither of which we want to change. For those reasons, a cryptographic hash function needs to be highly collision resistant. The drawback of collision resistant algorithms over simpler hashing functions is the number of operations needed for the calculation, while current hashing functions are performant and secure, they still take some time for much data \cite{crypto}.

One \gls{hash} is called \gls{sha256}. \gls{sha256} is a standard published by the \gls{nist}. It creates a 256-bit output and has not yet been successfully attacked \cite{sha, crypto}. This hashing algorithm is used in the implementation of the \gls{hids} to make sure that the configuration file has not changed between multiple runs.


\subsection{HIDS}
\label{sec:def:hids}

A \gls{hids} works by detecting changes on the local host. It does that by looking at files, processes, configuration, logs, or other indicators. In this thesis, I focus on \gls{fim}. It is important to note that the other origins mentioned are also a valuable source of information. Any of those might have \gls{ioc}. Especially running processes and the configuration can hold relevant data. However, this implementation only covers file-based information due to the lack of resources.


In a \gls{hids} that uses \gls{fim}, all the valuable information comes from unexpected changes to the \gls{fs}. Many hosts do not have any changes on the \gls{fs} except for software updates. Additionally, the \gls{fs} usually contains too much information to cover in a handwritten configuration. Thus, the easiest approach is to compare the current state of the \gls{fs} to previous ones. If it has changed significantly or in an unexpected way, something foul might have happened or might still be happening. The negative side effect of this approach is the false positives that come from legitimately changing the host. Those legitimate changes could be new versions of the webpage that is running, newly updated configuration or updated ssh keypairs. However, those changes are scheduled, and the alerts can then be quickly checked and acknowledged.

A good \gls{hids} should be able to handle such valid changes without change on the configuration or other changes on the system. Additionally, it should be able to find intrusions reliably and in a timely fashion. Maybe, it is too late if the \gls{hids} finds an intrusion a week after the infection. To detect changes on the files reliably prominent \gls{hids} calculate a hash of the files and compares that hash to previous runs. If a \gls{hash} is used, each file always generates a unique hash which can not be forged. The main drawback of using \glspl{hash} is that they take a long time to compute for significant amounts of data. Some implementations thus offer to use \glspl{nonhash}. The obvious drawback of this approach is that a collision can be generated and such a file can be altered or replaced without the \gls{hids} noticing.

In the following sections, I present three of the most used \gls{hids}, \Gls{tripwire}, \Gls{aide}, and \Gls{samhain}. They built the main competition of the integration developed in this thesis.

\subsubsection{Tripwire}
\label{sec:tripwire}

In 1992 the first \gls{hids} named \gls{tripwire} was created and publicly released as a free tool. In 1997 the creator of \gls{tripwire} then created the company \Gls{tripwire} Inc. and bought the naming rights for \gls{tripwire}. The free version was monetized, and they released new versions of \gls{tripwire} \cite{Tripwire:Impl, Tripwire:company}. Now they mostly market to enterprise and industrial customer and have multiple products for securing computer systems and networks. \cite{tripwire}

From the documentation that they provide to potential customers, it is unclear how their product works exactly. This approach makes sense because this way, they can keep competitors at bay. However, they have some information about how their product works to attract new customers. They claim to find changed data in real time. Additionally, they seem to be able to know what part of the documents changed. They then use another product of theirs called ChangeIQ to determine if the change was malicious or not. \cite{tripwire:fim:datasheet, tripwire:true:intent}

For this approach to work, they need to create a big database with the content of each file. Then \gls{tripwire} probably is hooked into the kernel to get notified about changes to the \gls{fs} in real time. This approach has benefits over hashing and over checking attributes because it is faster than the former and more reliable than the latter. However, for that to work, the system needs to be running at all times. Additionally, not every \gls{os} supports this kind of kernel level information. I could not find out how they find intrusions should they ever be shut down and restarted at a later time or on \gls{os} on which this information is not available. 

What has to be said, \gls{tripwire} does not directly compete with my solution. \Gls{tripwire} is a commercial product with integration into monitoring and fancy user interfaces. They have much financial incentive to continuously improve their product and have high integration with an extensive suite of security tools aimed at big cooperations. I encourage people working for this type of company to have a look at \gls{tripwire}. However, because it mainly focuses on those big corporate customers, it might be too expensive for smaller teams. My \gls{hids} is more focused on smaller companies, teams within large companies, and participants of the \gls{opensource} community and is entirely free.

Even though the direct comparison might not even be relevant, my system has some advantages that I want to list here. It does not need a large amount of storage. For tripwire to work, it needs to store files entirely or partially to compare them to previous states accurately. This file storage might not be a problem for small files, but it does take much space if many machines are scanned or the amount of small files is extensive. My system only saves \gls{metadata}, which is considerably less space.

Additionally, my system does not need to run at all times. This approach reduces the load, which is especially essential on high-performance machines, particularly if they have a lot of read-write operations. One final advantage is the fact that my system is \gls{opensource}. With that, any potentially interested party can look at the source code and see if the product is appealing to them. Additionally, they can check the source code for security issues and can find out exactly how the system works. This information is not available for \gls{tripwire} customer. They have to trust the company behind it.


\subsubsection{AIDE}
\label{sec:aide}

\Gls{aide} is an \gls{opensource} alternative to \gls{tripwire}. It was created after \gls{tripwire} went commercial. As \gls{aide} is \gls{opensource}, it is easier to find information about how it works, which I detailed below. \cite{aide:totherescue, aide:github}

When \gls{aide} is first executed, it generates a reference database. This database contains all the information for each file that is within a path of interest. Depending on the config it contains more or less information, including \glspl{hash}. Each subsequent execution then compares the files found to this initially created database. The database needs to be updated manually if the system has a scheduled change. It generates a log of all the changes and alerts all of the changes at once. Both the configuration and the database are usually stored on the file system that is scanned. \cite{aide, aide:doc}

\Gls{aide} can compare multiple \glspl{hash} and \glspl{nonhash} and some \gls{fs} \gls{metadata}. It runs on many modern Unix system \cite{aide} and uses \gls{pgp} keys to sign their releases. It has a strong community behind it, but only offers a \gls{cli} and has no fancy user interface and is written in the programming language C. \cite{aide:github} 

\Gls{aide} also features a rich configuration. It is based on rules. There are some preconfigured rules for each hashing function and for the checked \gls{fs} \gls{metadata}. Those rules can then be combined in new rules to create custom settings. Finally, the configuration offers the possibility to specify paths which should be scanned with the rules that should be applied. This way, different directories can be scanned with different settings. \cite{aide:conf}

\subsubsection{Samhain}
\label{sec:samhain}

\Gls{samhain} is another \gls{opensource} \gls{hids}. In comparison to \gls{aide} it has a company backing it which offers support for customers. However, other than \gls{tripwire} it is not necessary to use their services to run \gls{samhain}. For that, they have a lot of publicly available documentation and community support. \cite{samhain:company}

\Gls{samhain} has a big focus on centralized management of the \gls{hids}. All hosts run the system and report their findings to a single instance which is used for management and monitoring. The hosts also get the configuration from the server. For scanning, it has two modes. Firstly, it can scan the system similarly to \gls{aide} by going through all the files and mostly generating hashes. However, it can also create a hook in the Linux kernel to find changes live. It uses both approaches to find intrusions. Contrary to \gls{aide} \gls{samhain} offers more functionality than \gls{fim} and can check open ports, processes and more. 

Compared to my system, it can leverage the kernel integration to find intrusions faster. However, it still relies on hashing for scanning parts of the \gls{fs}, especially on non-Linux systems. This part of \gls{samhain} is improved by my system. Overall, \gls{samhain} offers more functionality than \gls{fim}, which my system currently does not. For kernel integration, it needs to be running at all times. As already discussed in the tripwire section, this can be a drawback.

\subsection{NIDS}
\label{sec:def:nids}

Compared to a \gls{hids}, a \gls{nids} can be used to detect intrusions from multiple hosts at the same time. \gls{nids} are used to analyze anomalies on the network. An \gls{anomaly} might be an unexpected session from a server which usually doesn't communicate with the internet or unusually large or frequent traffic. Certain intrusions can be tracked by analyzing the content of the packages. Some intrusions are easier to detect on the network. Especially such ones that extract much data or are very aggressive at spreading within a network. The main advantage here is that multiple sources can be combined on the network level. Another advantage this approach has is that current \gls{malware} almost always contains some network communication \cite{Malware:Behaviour,nids}.

However, a \gls{nids} doesn't only have advantages. Certain kinds of attacks can only be detected on the network with much difficulty. For a corrupted file upload cannot be detected by only checking the network. For best results, both types of \gls{ids} can be used in conjunction to detect attacks. 

In this thesis \gls{nids} are not in the center. The implementation doesn't look at network traffic or mimic other \gls{nids} functionality. However, as mentioned above, I highly encourage everyone to use an appropriate \gls{nids} to improve the chance to find an intrusion. 

\subsection{The Sleuth Kit}
\label{sec:tsk}

\gls{tsk} is an \gls{opensource} toolkit used to investigate disk images. It is based on the coroner's toolkit \cite{tct} and contains multiple command line interfaces and an \gls{api} for various purposes. \cite{tsk, tsk:about} It is mainly written in C, runs on Linux, OsX, Windows and more and can be used to analyze many different \glspl{fs}. It is heavily used in forensic investigations to find deleted or corrupted files and for other information gatherings on a disk image.

\subsubsection{fls}
\label{sec:fls}

One of those command line tools is fls. It can be used to access directories, files, and the attributes of each. With it, the directories can be displayed recursively, and for each file, the attributes can be printed to the console. \cite{tsk:fls} This tool plays a central part in this thesis as it is used to recover all the files and attributes to detect intrusions. On how exactly it is used, please refer to section. \ref{sec:Scanner}

\subsection{Python}
\label{sec:python}

Python was the programming language of choice for this product. Python offers many advantages over other languages. Some of those are:

\begin{itemize}
    \item Platform independent
    \item Good community support in the forensic community
    \item Active library for \gls{tsk}
    \item Small overhead
    \item Easy for prototyping
    \item Easy to read
\end{itemize}

This decision was not made very lightly. Other programming languages were considered. For more information on those refer to section \ref{sec:decisions:language}.

\subsubsection{pytsk3}
\label{sec:pytsk3}

Pytsk3 is the library mentioned above that creates bindings for python to the \gls{api} of \gls{tsk}. As \gls{tsk} this library is \gls{opensource} and is still active. It is hosted on \gls{github} and offers most of the functionality of the \gls{tsk} \gls{api}. Pytsk3 is used extensively in the scanner part of the implementation. For further information refer to section \ref{sec:Scanner}.

\section{Proposed solution}
\label{sec:hids}

The main principle behind a \gls{hids} is the detection of changed files. As already discussed in section \ref{sec:def:hids}, most tools rely on the calculation of hashes. This is generally a good approach since changes can be found very reliably; however, as already mentioned, it can drastically hinder the performance of the \gls{hids}. Sadly the actual performance lost cannot be clearly stated, as it heavily depends on what hardware is in use. However, considering the computational overhead of calculating a \gls{hash} it is clear that the time it takes grows with bigger \glspl{fs}. As \gls{storagemedia} has grown from a \gls{mb} to \gls{tb} so has the requirement to store more data. Creating hashes over such big systems is not viable as it can take a long time to create a hash of big amounts of data. \cite{hash:slow, hash:veryslow, hash:speed}

\subsection{Time issues}

In this thesis, I propose a different approach. Ignore the hashing and the file content and focus on the \gls{fs} \gls{metadata} instead. This approach should be sufficient to find intrusions and thanks to not creating hashes, the \gls{hids} can execute more often. If a conventional \gls{hids} might take several days to complete a run on a \gls{fs} with multiple \gls{tb} of data, the proposed approach would take some minutes to hours. It could then be run multiple times a day and find new and changed files within a short amount of time. It is possible that the \gls{hids} misses some changes if the attacker can change all the relevant \gls{metadata} before the system rechecks the same file, but this is a risk that has to be taken to gain the opportunity to scan large \gls{fs}. It is also possible to scan highly critical sections of a system with a traditional \gls{hids} and the rest with the proposed solution. This way, one has both advantages. The whole system can be checked in a timely fashion with the proposed solution, and for certain smaller parts of the system, a general \gls{hids} can detect changes by using strong \glspl{hash}.

This solution uses \gls{tsk} via \gls{pytsk} to extract the \gls{fs} \gls{metadata}. The main advantage that this gives is the interoperability with different \glspl{fs}. Additionally, by directly accessing the attributes from the \gls{fs}, no \gls{metadata} is changed. The files themselves are never touched. Furthermore, it can also be used to get the files and attributes of an image that has been extracted or of a virtual machine. 

\subsection{Investigation capabilities}
\label{sec:investigation:capabilities}
Another improvement upon existing \gls{hids} proposed is that investigation capabilities need to be built in from the start. Traditional \gls{hids} are good at finding intrusions and alerting them. However, they don't offer support for the investigation of the incident. They don't have information beyond what was configured. This lack of information makes sense is dangerous \gls{hash}. If something was missed in the configuration, an investigator can't use the output of those systems to gain further knowledge. They only have the alerts and nothing further.

This solution stores all the available \gls{metadata} for each scanned file for each run. This way, an investigator can use this output to gain more information about how the attacker proceeded. He can look at the changes of permissions, modification of files, even if the alerts might have been ignored. It is also possible to generate a timeline out of this data to form an extended view on what happened when on the \gls{fs}. 

\subsection{Flexibility}

\Gls{aide} works by comparing runs to one first execution. This is practical as it detects one intrusion multiple times. However, it generates many messages, and users are then less likely to take them seriously. Additionally to that, after a legitimate change to the system, \gls{aide} always generates alerts until the initial run is reset. This can lead to undetected intrusions a short time after an upgrade, which might also be the most critical window for an intrusion because the update might have created a security risk. Thus, it is possible that an intruder can gain access shortly after an update which is alerted by \gls{aide}, but ignored because of the false positives.

The proposed system does not have this issue, at least not as strongly. As all the data gets collected for each run, it makes sense to compare each run to the previous. This has the benefit of legitimate changes being adopted into the accepted one run after it has been finished. This results in overall fewer messages, which increases the importance of each. If it is required that the system should always compare against one specific run, this could also clearly be done. The system would need to be configured to not check against the latest, but a specific, predefined run. Only the configuration would then need to be changed to update this predefined run. 

\section{Scope}

In this project I create a \gls{hids} which uses \gls{tsk} to detect changes. It covers the three main changes discussed in section \ref{sec:hids}. This service uses a \gls{sql} database to store the executed runs. It includes user documentation and this thesis documentation. 

Out of scope is the creation of the timeline mentioned in section \ref{sec:investigation:capabilities}. Also, out of scope are extensive alerting functionality and extensive example configurations for commonly used \gls{os} and tools. Furthermore, any big data analysis of the runs, while very interesting, are also out of scope.

